---
params: 
    input_la_choice: Durham
    input_phase_choice: Primary
    input_quality_chart: 'Ofsted'
header-includes:
   - \usepackage{graphics}
   - \usepackage{expl3}
   - \usepackage{xparse}
   - \usepackage{tcolorbox}
   - \usepackage{amsmath,amsfonts,amsthm}
   - \usepackage{wrapfig}
   - \usepackage{helvet}
   - \usepackage{sectsty}
   - \usepackage{fancyhdr}
   - \usepackage{xpatch}
   - \pagestyle{fancy}
   - \definecolor{gssmidblue}{RGB}{32, 115, 188}
   - \definecolor{dfeheadingblue}{RGB}{16, 79, 117}
   - \renewcommand{\familydefault}{\sfdefault}
   - \allsectionsfont{\color{dfeheadingblue}}
   - \fancyhead[C,L,R]{}
   - \fancyfoot[R]{\nouppercase{\emph{\rightmark}}}
   - \fancyfoot[L]{\nouppercase{\emph{\leftmark}}}
   - \fancyfoot[C] {}
   - \renewcommand{\headrulewidth}{0pt}
   - \renewcommand{\footrulewidth}{2pt}
   - \futurelet\TMPfootrule\def\footrule{{\color{gssmidblue}\TMPfootrule}}
output:
  pdf_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::knit_hooks$set(crop = knitr::hook_pdfcrop)
source("global.R")
library(webshot)
library(kableExtra)

```


\includegraphics[width=0.25\linewidth]{"images/Department_for_Education.png"}
\vspace{1.6cm}


# `r params$input_phase_choice` School Places Scorecard for `r params$input_la_choice`

## Introduction

This scorecard presents school places benchmarking figures for `r params$input_la_choice` covering quantity of places available, places allocated by preference, places allocated by school quality and funding provided for new places.

```{r basic need funding, echo = FALSE}

# Take data, get total funding and divide by billion if it's England, million if it's not
    total_funding <- scorecards_data %>%
      filter(LA_name == params$input_la_choice) %>%
      select(Funding) %>%
      mutate(
        Funding =
          ifelse(params$input_la_choice == "England", roundFiveUp(Funding / 1000000000, 2),
            roundFiveUp(Funding / 1000000, 0)
          )
      ) %>%
      as.numeric()

# Take filtered data, search for growth rate, pull the value and tidy the number up
  live_scorecard_data <- scorecards_data_pivot %>% filter(
      LA_name == params$input_la_choice,
      Phase == params$input_phase_choice
    )
  
    growth_perc <- live_scorecard_data %>%
      filter(name == "Bangro") %>%
      pull(value) %>%
      roundFiveUp(., 2) * 100

    # PDisplay growth in pupil numbers
   textTopLineGrowth <- paste0("Growth in ", str_to_lower(params$input_phase_choice), " pupil numbers 2009/10 to ", plan_year, ": ", growth_perc, "%")
   
   dfTopLine <- data.frame(funding=ifelse(params$input_la_choice== "England",
                                          paste0(total_funding, "billion"),
                                          paste0(total_funding, "million")),
                           empty='',
                           growth=paste0(growth_perc,"%"))
   colnames(dfTopLine) <- c(paste0("Total primary and secondary basic need funding ",funding_year),' ',
                            paste0("Growth in",str_to_lower(params$input_phase_choice), "pupil numbers 2009/10 ",plan_year)
   )

```


\makebox[1.00\linewidth]{
\centering


\begin{tcolorbox}[colback=gssmidblue, 
 leftright skip=0.1cm,
 coltext=white, 
 halign=left, 
 fontupper={\Huge \bfseries},
 fontlower={\large \bfseries},
 sharp corners, 
 colframe=gssmidblue,
 width=0.49\linewidth,
 boxrule=0pt 
 ]
£`r ifelse(params$input_la_choice== "England",paste0(total_funding, "billion"),paste0(total_funding, "million"))`
\tcblower
Total primary and secondary basic need funding `r funding_year`
\end{tcolorbox}


\begin{tcolorbox}[colback=gssmidblue, 
 leftright skip=0.1cm,
 coltext=white, 
 halign=left, 
 fontupper={\Huge \bfseries},
 fontlower={\large \bfseries},
 sharp corners, 
 colframe=gssmidblue,
 width=0.49\linewidth,
 boxrule=0pt 
 ]
`r growth_perc`\%
\tcblower
Growth in `r str_to_lower(params$input_phase_choice)`, pupil numbers 2009/10 to `r plan_year`
\end{tcolorbox}
}

\newpage
  
## Quantity

### Places created since 2009-10, places planned to 2021-22 and estimated place pressure in 2021-22

A local authority can have both ‘spare places’ and ‘additional places needed’ due to localised or specific year group demand

```{r quantity, echo = FALSE}
   
  ## Estimated places need

    # Take filtered data, search for growth rate, pull the value and tidy the number up
    additional_places_perc <- live_scorecard_data %>%
      filter(name == "QuanRP") %>%
      pull(value)

    # Display value
           x <- paste0("Estimated additional ", str_to_lower(params$input_phase_choice), " places to meet demand in ", plan_year, ": ", scales::comma(additional_places_perc))
     
  

  ## Estimated spare places

    # Take filtered data, search for growth rate, pull the value and tidy the number up
    spare_places_per <- live_scorecard_data %>%
      filter(name == "QuanSu") %>%
      pull(value) %>%
      roundFiveUp(., 2) * 100
```            

\makebox[1.0\linewidth]{
\centering
\begin{tcolorbox}[colback=gssmidblue, 
 leftright skip=0.1cm,
 coltext=white, 
 halign=left, 
 fontupper={\Huge \bfseries},
 fontlower={\large \bfseries},
 sharp corners, 
 colframe=gssmidblue,
 width=0.49\linewidth,
 boxrule=0pt 
 ]
`r scales::comma(additional_places_perc)`
\tcblower
Estimated additional `r str_to_lower(params$input_phase_choice)` places to meet demand in `r plan_year` 
\end{tcolorbox}
\begin{tcolorbox}[colback=gssmidblue, 
 leftright skip=0.1cm,
 coltext=white, 
 halign=left, 
 fontupper={\Huge \bfseries},
 fontlower={\large \bfseries},
 sharp corners, 
 colframe=gssmidblue,
 width=0.49\linewidth,
 boxrule=0pt 
 ]
`r spare_places_per`\%
\tcblower
Estimated percentage of spare `r str_to_lower(params$input_phase_choice)` places in `r plan_year`
\end{tcolorbox}
}

``` {r quantity_b, echo = FALSE, fig.show="hold", out.width = "100%",crop=TRUE}
 live_scorecard_data <- scorecards_data_pivot %>% filter(
      LA_name == params$input_la_choice,
      Phase == params$input_phase_choice)

    places_chart_data <- live_scorecard_data %>%
      filter(name %in% c("QuanIn", "QuanPP", "QuanRP")) %>%
      select(LA_name, name, value) %>%
      pivot_wider()  
   
quality_places <- plot_ly(
      places_chart_data,
      width = 420, height = 560,
      x = ~LA_name, y = ~QuanIn,
      marker = list(color = c("#12436D")),
      type = "bar", name = paste0("Total places created between 2009/10 and ", this_year),
      text = ~ scales::comma(QuanIn), textposition = "inside", textfont = list(color = "#FFF")
    ) %>%
      add_trace(y = ~QuanPP, marker = list(color = c("#F46A25")), name = paste0("New places already planned for delivery between ", this_year, " and ", plan_year), text = ~ scales::comma(QuanPP), textposition = "inside") %>%
      add_trace(y = ~QuanRP, marker = list(color = c("#801650")), name = paste0("Estimated additional places still needed to meet demand in ", plan_year), text = ~ scales::comma(QuanRP), textposition = "inside") %>%
      layout(
        yaxis = list(title = ""),
        xaxis = list(title = ""),
        barmode = "stack",
        uniformtext = list(minsize = 12, mode = "hide"),
        legend = list(orientation = "h"),
        title = list(
          text = "Chart showing total places created, new places planned for delivery and estimated additional places needed to meet demand, by Local Authority compared to England",
          font = list(color = "#ffffff")
        )
      ) %>%
      config(displayModeBar = FALSE)
      htmlwidgets::saveWidget(widget = quality_places, file = "hc.html")
      webshot(url = "hc.html", file = "hc.png", delay = 1, zoom = 4, vheight = 500)
```

\newpage

## Preference

### Proportion of applicants who received an offer of one of their top three preference schools for September 2019 entry

```{r preference_text, echo = FALSE, message=FALSE}

live_scorecard_data_all_la <- scorecards_data_pivot %>% filter(Phase == params$input_phase_choice)


    # Take filtered data, search for growth rate, pull the value and tidy the number up
    PrefT3_E <- live_scorecard_data_all_la %>%
      filter(name == "PrefT3") %>%
      filter(LA_name == "England") %>%
      pull(value) %>%
      roundFiveUp(., 2)

      # Take filtered data, search for growth rate, pull the value and tidy the number up
    PrefT3 <- live_scorecard_data %>%
      filter(name == "PrefT3") %>%
      pull(value) %>%
      roundFiveUp(., 2)
```

\makebox[1.0\linewidth]{
\centering
\begin{tcolorbox}[colback=gssmidblue, 
 leftright skip=0.1cm,
 coltext=white, 
 halign=left, 
 fontupper={\Huge \bfseries},
 fontlower={\large \bfseries},
 sharp corners, 
 colframe=gssmidblue,
 width=0.49\linewidth,
 boxrule=0pt 
 ]
`r PrefT3_E`\%
\tcblower
Percentage of applicants who recieved an offer of one of their top three preferred `r str_to_lower(params$input_phase_choice)` schools in England
\end{tcolorbox}
``` {r echo = FALSE, warning = FALSE, message=FALSE, results='asis'}
if (params$input_la_choice != "England") {
cat(paste("\\begin{tcolorbox}[colback=gssmidblue, 
 leftright skip=0.1cm,
 coltext=white, 
 halign=left, 
 fontupper={\\Huge \\bfseries},
 fontlower={\\large \\bfseries},
 sharp corners, 
 colframe=gssmidblue,
 width=0.49\\linewidth,
 boxrule=0pt 
 ]
",PrefT3,"\\%
\\tcblower
Percentage of applicants who recieved an offer of one of their top three preferred ", str_to_lower(params$input_phase_choice)," places in ",plan_year," in ",params$input_la_choice,"
\\end{tcolorbox}
}
"))
    } else {
paste("        
}
")
      }
```

```{r preference_barchart, echo = FALSE, message=FALSE, out.width="100%", crop=TRUE}
# Scorecard data, filtered on user input AND including England as a comparison
  live_scorecard_data_england_comp <- scorecards_data_pivot %>%
      filter(
        LA_name %in% c(params$input_la_choice, "England"),
        Phase == params$input_phase_choice
      ) %>%
      mutate(
        LA_name = as.factor(LA_name),
        # This step just makes sure that the LA is FIRST when it comes to plots/tables
        LA_name = relevel(LA_name, "England"),
        LA_name = factor(LA_name, levels = rev(levels(LA_name)))
      )
  
    # reshape the data so it plots neatly!
    preference_data <- live_scorecard_data_england_comp %>%
      # select only preference values
      filter(name %in% c("Pref1", "Pref2", "Pref3")) %>%
      # Create ratings out of the names
      mutate(rating = case_when(
        str_detect(name, "1") ~ "First choice",
        str_detect(name, "2") ~ "Second choice",
        str_detect(name, "3") ~ "Third choice"
      ))

    # Get % not getting 1st 2nd or 3rd preference
    preference_data_sum <- preference_data %>%
      group_by(LA_name, LANumber, Phase) %>%
      summarise(value = 100 - sum(value)) %>%
      mutate(rating = "Other")


    preference_data <- preference_data %>%
      select(-name) %>%
      bind_rows(preference_data_sum) %>%
      # sort levels out so plots in correct order
      mutate(
        rating = factor(rating, levels = c("First choice", "Second choice", "Third choice", "Other")),
        # Neaten up percs
        value = as.numeric(roundFiveUp(value, 1)),
        value_label = if_else(value > 3, paste0(value, "%"), NA_character_)
      )


# create charts
 preference_p <- preference_data %>%
      ggplot(aes(
        y = value, x = "",
        fill = factor(rating),
        text = paste(rating, ": ", value, "%")
      )) +
      geom_bar(stat = "identity", position = position_fill(reverse = TRUE)) +
      coord_flip() +
      facet_wrap(~LA_name, nrow = 2) +
      geom_text(aes(label = value_label), colour = "#ffffff", size = 4, 
                position = position_fill(reverse = TRUE, vjust = 0.5)) +
      labs(x = "", y = "") +
      guides(fill = guide_legend(title = "")) +
      scale_fill_manual(values = dfe_colours) +
      theme_minimal() +
      theme(
        legend.position = "bottom",
        text = element_text(size = 14, family = "Arial")
      )
    ggplotly(preference_p,
             width=960,height=480
         ) %>%
      layout(
        uniformtext = list(minsize = 12, mode = "hide"),
        xaxis = list(showticklabels = FALSE),
        legend = list(
          orientation = "h",
          y = -0.1, x = 0.
                 ),
        title = list(
          font = list(color = "#ffffff")
        )
      ) %>%
      config(displayModeBar = FALSE)
  
```

\newpage
## Quality

### Quality of places created between 2017-18 and 2018-19, based on `r params$input_quality_chart`:

```{r quality, echo = FALSE}
  # Change name of what "better than average" is depending on chart choice:
  school_description <- 
    if (params$input_quality_chart == "Ofsted") {
      "good and outstanding "
    } else {
      "well above and above average "
    }


  # Calculate LA % depending on chart choice:
  LA_comp <- 
    if (params$input_quality_chart == "Ofsted") {
      live_scorecard_data %>%
        filter(name == "QualProp") %>%
        pull(value) %>%
        roundFiveUp(., 2) * 100
    } else if (params$input_quality_chart == "Progress 8") {
      live_scorecard_data %>%
        filter(name == "Qual_KS4_Prop") %>%
        pull(value) %>%
        roundFiveUp(., 2) * 100
    } else if (params$input_quality_chart == "Reading Progress") {
      live_scorecard_data %>%
        filter(name == "Qual_KS2Read_Prop") %>%
        pull(value) %>%
        roundFiveUp(., 2) * 100
    } else if (params$input_quality_chart== "Maths Progress") {
      live_scorecard_data %>%
        filter(name == "Qual_KS2Mat_Prop") %>%
        pull(value) %>%
        roundFiveUp(., 2) * 100
    }
  
             # Scorecard data for ALL LAs, filtered only on phase choice
live_scorecard_data_all_la <- scorecards_data_pivot %>% filter(Phase == params$input_phase_choice)
 
        
  # Calculate England comparator depending on chart choice:
  england_comp <- 
    if (params$input_quality_chart == "Ofsted") {
      numerator <- live_scorecard_data_all_la %>%
        filter(LA_name == "England" &
          name %in% c("Qual1_N", "Qual2_N")) %>%
        summarise(sum(value)) %>%
        as.numeric()

      denominator <- live_scorecard_data_all_la %>%
        filter(LA_name == "England" &
          name %in% c("Qual1_N", "Qual2_N", "Qual3_N", "Qual4_N")) %>%
        summarise(sum(value)) %>%
        as.numeric()

      # calculate percentage
      roundFiveUp(numerator / denominator * 100, 1)
    } else if (params$input_quality_chart == "Progress 8") {
      numerator <- live_scorecard_data_all_la %>%
        filter(LA_name == "England" &
          name %in% c("KS4_WAA_N", "KS4_AA_N")) %>%
        summarise(sum(value)) %>%
        as.numeric()

      denominator <- live_scorecard_data_all_la %>%
        filter(LA_name == "England" &
          name %in% c("KS4_WAA_N", "KS4_AA_N", "KS4_A_N", "KS4_BA_N", "KS4_WBA_N")) %>%
        summarise(sum(value)) %>%
        as.numeric()

      # calculate percentage
      roundFiveUp(numerator / denominator * 100, 1)
    } else if (params$input_quality_chart == "Reading Progress") {
      numerator <- live_scorecard_data_all_la %>%
        filter(LA_name == "England" &
          name %in% c("KS2Read_WAA_N", "KS2Read_AA_N")) %>%
        summarise(sum(value)) %>%
        as.numeric()

      denominator <- live_scorecard_data_all_la %>%
        filter(LA_name == "England" &
          name %in% c("KS2Read_WAA_N", "KS2Read_AA_N", "KS2Read_A_N", "KS2Read_BA_N", "KS2Read_WBA_N")) %>%
        summarise(sum(value)) %>%
        as.numeric()

      # calculate percentage
      roundFiveUp(numerator / denominator * 100, 1)
    } else if (params$input_quality_chart == "Maths Progress") {
      numerator <- live_scorecard_data_all_la %>%
        filter(LA_name == "England" &
          name %in% c("KS2Mat_WAA_N", "KS2Mat_AA_N")) %>%
        summarise(sum(value)) %>%
        as.numeric()

      denominator <- live_scorecard_data_all_la %>%
        filter(LA_name == "England" &
          name %in% c("KS2Mat_WAA_N", "KS2Mat_AA_N", "KS2Mat_A_N", "KS2Mat_BA_N", "KS2Mat_WBA_N")) %>%
        summarise(sum(value)) %>%
        as.numeric()

      # calculate percentage
      roundFiveUp(numerator / denominator * 100, 1)
    }

            # Calculate % ranking depending on chart choice:
  LA_ranking <- 
    if (params$input_quality_chart == "Ofsted") {
      live_scorecard_data %>%
        filter(name == "QualPropranks") %>%
        pull(value)
    } else if (params$input_quality_chart == "Progress 8") {
      live_scorecard_data %>%
        filter(name == "Qual_KS4_Propranks") %>%
        pull(value)
    } else if (params$input_quality_chart == "Reading Progress") {
      live_scorecard_data %>%
        filter(name == "Qual_KS2Read_Propranks") %>%
        pull(value)
    } else if (params$input_quality_chart == "Maths Progress") {
      live_scorecard_data %>%
        filter(name == "Qual_KS2Mat_Propranks") %>%
        pull(value)
    }

    # Calculate ranking denominator depending on chart choice:
  LA_denom <- 
    if (params$input_quality_chart == "Ofsted") {
      live_scorecard_data_all_la %>%
        filter(name == "QualPropranks" & !is.na(value)) %>%
        nrow()
    } else if (params$input_quality_chart == "Progress 8") {
      live_scorecard_data_all_la %>%
        filter(name == "Qual_KS4_Propranks" & !is.na(value)) %>%
        nrow()
    } else if (params$input_quality_chart == "Reading Progress") {
      live_scorecard_data_all_la %>%
        filter(name == "Qual_KS2Read_Propranks" & !is.na(value)) %>%
        nrow()
    } else if (params$input_quality_chart == "Maths Progress") {
      live_scorecard_data_all_la %>%
        filter(name == "Qual_KS2Mat_Propranks" & !is.na(value)) %>%
        nrow()
    }
```           

\makebox[1.0\linewidth]{
\centering
\begin{tcolorbox}[colback=gssmidblue, 
 leftright skip=0.1cm,
 coltext=white, 
 halign=left, 
 fontupper={\Huge \bfseries},
 fontlower={\large \bfseries},
 sharp corners, 
 colframe=gssmidblue,
 `r ifelse(params$input_la_choice != "England","width=0.32\\linewidth","width=0.49\\linewidth")`,
 boxrule=0pt 
 ]
`r england_comp`\%
\tcblower
Percentage of new places created in `r paste(school_description, str_to_lower(params$input_phase_choice))` schools in England
\end{tcolorbox}
``` {r echo = FALSE, warning = FALSE, message=FALSE, results='asis'}
    if (params$input_la_choice != "England") {
cat(paste("\\begin{tcolorbox}[colback=gssmidblue, 
 leftright skip=0.1cm,
 coltext=white, 
 halign=left, 
 fontupper={\\Huge \\bfseries},
 fontlower={\\large \\bfseries},
 sharp corners, 
 colframe=gssmidblue,
 width=0.32\\linewidth,
 boxrule=0pt 
 ]
",LA_comp,"\\%
\\tcblower
Percentage of new places created in ", school_description, str_to_lower(params$input_phase_choice), " schools in ", params$input_la_choice,"
\\end{tcolorbox}
")
)
cat(paste("\\begin{tcolorbox}[colback=gssmidblue, 
 leftright skip=0.1cm,
 coltext=white, 
 halign=left, 
 fontupper={\\Huge \\bfseries},
 fontlower={\\large \\bfseries},
 sharp corners, 
 colframe=gssmidblue,
 width=0.32\\linewidth,
 boxrule=0pt 
 ]
",LA_ranking,"
\\tcblower
LA Rank out of ", LA_denom, " LAs that created new places between ", last_year, " and ", this_year, " (ranks can be tied)
\\end{tcolorbox}")
)
      }
```
}


``` {r quality_chart, out.width="92%", crop=TRUE}           
# Scorecard data, filtered on user input AND including England as a comparison
  live_scorecard_data_england_comp <- scorecards_data_pivot %>%
      filter(
        LA_name %in% c(params$input_la_choice, "England"),
        Phase == params$input_phase_choice
      ) %>%
      mutate(
        LA_name = as.factor(LA_name),
        # This step just makes sure that the LA is FIRST when it comes to plots/tables
        LA_name = relevel(LA_name, "England"),
        LA_name = factor(LA_name, levels = rev(levels(LA_name)))
      )
  
 # Bar chart comparison - Ofsted

    # reshape the data so it plots neatly!
 ofsted_data <- live_scorecard_data_england_comp %>%
      # select only the ofsted values
      filter(name %in% c(
        "Qual1_N", "Qual2_N", "Qual3_N", "Qual4_N", "Qual0_N",
        "Qual1_E", "Qual2_E", "Qual3_E", "Qual4_E", "Qual0_E"
      )) %>%
      # Create groups for "new" and "existing" places based on names
      mutate(place_type = case_when(
        str_detect(name, "N") ~ "New",
        str_detect(name, "E") ~ "Existing"
      )) %>%
      # Create Ofsted ratings out of the names
      mutate(rating = case_when(
        str_detect(name, "1") ~ "Oustanding",
        str_detect(name, "2") ~ "Good",
        str_detect(name, "3") ~ "Requires Improvement",
        str_detect(name, "4") ~ "Inadequate",
        str_detect(name, "0") ~ "No rating"
      )) %>%
      # Create new variable called places, replace 0s with NAs so it plots neatly
      mutate(
        places = if_else(value == 0, NA_integer_, as.integer(roundFiveUp(value, 0)))
      ) %>%
      # Give NA for label if it's too small
      group_by(LA_name, place_type) %>%
      mutate(
        places_perc = places / sum(places, na.rm = TRUE),
        value_label = if_else(places_perc > 0.05, places, NA_integer_)
      )

    ofsted_p <- ofsted_data %>%
      filter(rating != "No rating") %>%
      ggplot(aes(
        y = value, x = place_type,
        fill = factor(rating, levels = c("Oustanding", "Good", "Requires Improvement", "Inadequate")),
        text = paste(rating, ": ", places, " places")
      )) +
      geom_bar(stat = "identity", position = position_fill(reverse = TRUE)) +
      coord_flip() +
      facet_wrap(~LA_name, nrow = 2) +
      geom_text(aes(label = scales::comma(value_label)), size = 4, colour = "#FFFFFF", position = position_fill(reverse = TRUE, vjust = 0.5)) +
      labs(x = "", y = "") +
      guides(fill = guide_legend(title = "")) +
      scale_fill_manual(values = dfe_colours) +
      theme_minimal() +
      theme(
        legend.position = "bottom",
        text = element_text(size = 14, family = "Arial")
      )

    if (params$input_la_choice == "England" & params$input_quality_chart == "Ofsted") {
      ofsted_no_rating <- ofsted_data %>%
        filter(rating == "No rating" & place_type == "New") %>%
        pull(places)
    } else if (params$input_quality_chart == "Ofsted") {
      ofsted_no_rating <- ofsted_data %>%
        filter(LA_name != "England" & rating == "No rating" & place_type == "New") %>%
        pull(places)
    }

    
    # Bar chart comparison - Progress 8

    # reshape the data so it plots neatly!
    progress_8_data <- live_scorecard_data_england_comp %>%
      # select only the progress 8 values
      filter(name %in% c(
        "KS4_WAA_N", "KS4_AA_N", "KS4_A_N", "KS4_BA_N", "KS4_WBA_N", "KS4_NR_N",
        "KS4_WAA_E", "KS4_AA_E", "KS4_A_E", "KS4_BA_E", "KS4_WBA_E", "KS4_NR_E"
      )) %>%
      # Create groups for "new" and "existing" places based on names
      mutate(place_type = case_when(
        str_detect(name, "N$") ~ "New",
        str_detect(name, "E$") ~ "Existing"
      )) %>%
      # Create Ofsted ratings out of the names
      mutate(rating = case_when(
        str_detect(name, "_WAA_") ~ "Well above average",
        str_detect(name, "_AA_") ~ "Above average",
        str_detect(name, "_A_") ~ "Average",
        str_detect(name, "_BA_") ~ "Below average",
        str_detect(name, "_WBA_") ~ "Well below average",
        str_detect(name, "NR") ~ "No rating"
      )) %>%
      mutate(rating = factor(rating, levels = c("Well above average", "Above average", "Average", "Below average", "Well below average", "No rating"))) %>%
      # Create new variable called places, replace 0s with NAs so it plots neatly
      mutate(places = if_else(value == 0, NA_integer_, as.integer(roundFiveUp(value, 0)))) %>%
      # Give NA for label if it's too small
      group_by(LA_name, place_type) %>%
      mutate(
        places_perc = places / sum(places, na.rm = TRUE),
        value_label = if_else(places_perc > 0.05, places, NA_integer_)
      )



    progress_8_p <- progress_8_data %>%
      filter(rating != "No rating") %>%
      ggplot(aes(
        y = value, x = place_type,
        text = paste(rating, ": ", places, " places"),
        group = rating,
        fill = rating
      )) +
      geom_bar(stat = "identity", position = position_fill(reverse = TRUE)) +
      coord_flip() +
      facet_wrap(~LA_name, nrow = 2) +
      geom_text(aes(label = scales::comma(value_label)), size = 4, colour = "#FFFFFF", position = position_fill(reverse = TRUE, vjust = 0.5)) +
      labs(x = "", y = "") +
      guides(fill = guide_legend(title = "")) +
      scale_fill_manual(values = dfe_colours) +
      theme_minimal() +
      theme(
        legend.position = "bottom",
        text = element_text(size = 14, family = "Arial")
      )

    if (params$input_la_choice == "England" & params$input_quality_chart == "Progress 8") {
      progress_8_no_rating <- progress_8_data %>%
        filter(rating == "No rating" & place_type == "New") %>%
        pull(places)
    } else if (params$input_quality_chart == "Progress 8") {
      progress_8_no_rating <- progress_8_data %>%
        filter(LA_name != "England" & rating == "No rating" & place_type == "New") %>%
        pull(places)
    }
 
        # Bar chart comparison - Progress Reading

    # reshape the data so it plots neatly!
    progress_reading_data <- live_scorecard_data_england_comp %>%
      # select only the reading values
      filter(name %in% c(
        "KS2Read_WAA_N", "KS2Read_AA_N", "KS2Read_A_N", "KS2Read_BA_N", "KS2Read_WBA_N", "KS2Read_NR_N",
        "KS2Read_WAA_E", "KS2Read_AA_E", "KS2Read_A_E", "KS2Read_BA_E", "KS2Read_WBA_E", "KS2Read_NR_E"
      )) %>%
      # Create groups for "new" and "existing" places based on names
      mutate(place_type = case_when(
        str_detect(name, "N$") ~ "New",
        str_detect(name, "E$") ~ "Existing"
      )) %>%
      # Create Ofsted ratings out of the names
      mutate(rating = case_when(
        str_detect(name, "_WAA_") ~ "Well above average",
        str_detect(name, "_AA_") ~ "Above average",
        str_detect(name, "_A_") ~ "Average",
        str_detect(name, "_BA_") ~ "Below average",
        str_detect(name, "_WBA_") ~ "Well below average",
        str_detect(name, "_NR_") ~ "No rating"
      )) %>%
      mutate(rating = factor(rating, levels = c("Well above average", "Above average", "Average", "Below average", "Well below average", "No rating"))) %>%
      # Create new variable called places, replace 0s with NAs so it plots neatly
      mutate(places = if_else(value == 0, NA_integer_, as.integer(roundFiveUp(value, 0)))) %>%
      # Give NA for label if it's too small
      group_by(LA_name, place_type) %>%
      mutate(
        places_perc = places / sum(places, na.rm = TRUE),
        value_label = if_else(places_perc > 0.05, places, NA_integer_)
      )


    progress_reading_p <- progress_reading_data %>%
      filter(rating != "No rating") %>%
      ggplot(aes(
        y = value, x = place_type,
        text = paste(rating, ": ", places, " places"),
        group = rating,
        fill = rating
      )) +
      geom_bar(stat = "identity", position = position_fill(reverse = TRUE)) +
      coord_flip() +
      facet_wrap(~LA_name, nrow = 2) +
      geom_text(aes(label = scales::comma(value_label)), size = 4, colour = "#FFFFFF", position = position_fill(reverse = TRUE, vjust = 0.5)) +
      labs(x = "", y = "") +
      guides(fill = guide_legend(title = "")) +
      scale_fill_manual(values = dfe_colours) +
      theme_minimal() +
      theme(
        legend.position = "bottom",
        text = element_text(size = 14, family = "Arial")
      )


    if (params$input_la_choice == "England" & params$input_quality_chart == "Reading Progress") {
      progress_reading_no_rating <- progress_reading_data %>%
        filter(rating == "No rating" & place_type == "New") %>%
        pull(places)
    } else if (params$input_quality_chart == "Reading Progress") {
      progress_reading_no_rating <- progress_reading_data %>%
        filter(LA_name != "England" & rating == "No rating" & place_type == "New") %>%
        pull(places)
    }

    # Bar chart comparison - Progress Maths

    # reshape the data so it plots neatly!
    progress_maths_data <- live_scorecard_data_england_comp %>%
      # select only the maths values
      filter(name %in% c(
        "KS2Mat_WAA_N", "KS2Mat_AA_N", "KS2Mat_A_N", "KS2Mat_BA_N", "KS2Mat_WBA_N", "KS2Mat_NR_N",
        "KS2Mat_WAA_E", "KS2Mat_AA_E", "KS2Mat_A_E", "KS2Mat_BA_E", "KS2Mat_WBA_E", "KS2Mat_NR_E"
      )) %>%
      # Create groups for "new" and "existing" places based on names
      mutate(place_type = case_when(
        str_detect(name, "N$") ~ "New",
        str_detect(name, "E$") ~ "Existing"
      )) %>%
      # Create Ofsted ratings out of the names
      mutate(rating = case_when(
        str_detect(name, "_WAA_") ~ "Well above average",
        str_detect(name, "_AA_") ~ "Above average",
        str_detect(name, "_A_") ~ "Average",
        str_detect(name, "_BA_") ~ "Below average",
        str_detect(name, "_WBA_") ~ "Well below average",
        str_detect(name, "NR") ~ "No rating"
      )) %>%
      mutate(rating = factor(rating, levels = c("Well above average", "Above average", "Average", "Below average", "Well below average", "No rating"))) %>%
      # Create new variable called places, replace 0s with NAs so it plots neatly
      mutate(places = if_else(value == 0, NA_integer_, as.integer(roundFiveUp(value, 0)))) %>%
      # Give NA for label if it's too small
      group_by(LA_name, place_type) %>%
      mutate(
        places_perc = places / sum(places, na.rm = TRUE),
        value_label = if_else(places_perc > 0.05, places, NA_integer_)
      )


    progress_maths_p <- progress_maths_data %>%
      filter(rating != "No rating") %>%
      ggplot(aes(
        y = value, x = place_type,
        text = paste(rating, ": ", places, " places"),
        group = rating,
        fill = rating
      )) +
      geom_bar(stat = "identity", position = position_fill(reverse = TRUE)) +
      coord_flip() +
      facet_wrap(~LA_name, nrow = 2) +
      geom_text(aes(label = scales::comma(value_label)), size = 4, colour = "#FFFFFF", position = position_fill(reverse = TRUE, vjust = 0.5)) +
      labs(x = "", y = "") +
      guides(fill = guide_legend(title = "")) +
      scale_fill_manual(values = dfe_colours) +
      theme_minimal() +
      theme(
        legend.position = "bottom",
        text = element_text(size = 14, family = "Arial")
      )

    if (params$input_la_choice == "England" & params$input_quality_chart == "Maths Progress") {
      progress_maths_no_rating <- progress_maths_data %>%
        filter(rating == "No rating" & place_type == "New") %>%
        pull(places)
    } else if (params$input_quality_chart == "Maths Progress") {
      progress_maths_no_rating <- progress_maths_data %>%
        filter(LA_name != "England" & rating == "No rating" & place_type == "New") %>%
        pull(places)
    }
    
  
    
    # Pick chart to plot based on user input
    if (params$input_quality_chart == "Ofsted") {
      quality_chart_choice <- ofsted_p
      no_rating <- ofsted_no_rating
    } else if (params$input_quality_chart == "Reading Progress") {
      quality_chart_choice <- progress_reading_p
      no_rating <- progress_reading_no_rating
    } else if (params$input_quality_chart == "Maths Progress") {
      quality_chart_choice <- progress_maths_p
      no_rating <- progress_maths_no_rating
    } else if (params$input_quality_chart == "Progress 8") {
      quality_chart_choice <- progress_8_p
      no_rating <- progress_8_no_rating
    }
    

    ggplotly(quality_chart_choice,
             width=900, height=480,
         ) %>%
      layout(
        xaxis = list(showticklabels = FALSE),
        legend = list(
          orientation = "h",
          y = -0.1, x = 0.2
                ),
        title = list(
          text = "Chart showing the quality of new and existing school places and estimated additional places, by Local Authority compared to England",
          font = list(color = "#ffffff")
        )
      ) %>%
      config(displayModeBar = FALSE)
    
```

New places with no rating =  `r scales::comma(no_rating)` 


## Cost


